---
title: "Machine Learning"
author: "Sachin Kekatpuray"
date: "Sunday, November 22, 2015"
output: html_document
---

##Summary

In this project I have utlized Random Forest algorithms as means to obtain a data-model towards predicting exercise classifications based on a certain number of predictors. The project is executed in following summarized steps:

1. Read training and test data.
2. Cleanup training data, remove unneccessary variables from the predictor list. Most the factor variables and variables having significant NAs. 
3. Partition training data into training and test data.  We shall use this test data for cross validation.
4. Generate model by applying random forest algorithm on training's train data. 
5. Run prediction and confusion matrix on training's test data.
6. Predict Test-data from step-1.

```{r, echo=FALSE,warning=FALSE}
library(caret)
library(kernlab)
library(stringr)
library(randomForest)
library(stats)
setwd("C:/dev/DataScience/MachineLearning/Proj")

```
###Read data
```{r, echo=TRUE}
training <- read.csv("pml-training.csv", header=T)
testing <- read.csv("pml-testing.csv", header=T)
```

###Cleanup data
Remove factor variables, and other variables having significant NAs.
```{r, echo=TRUE}
colCnt <- dim(training)[2] - 1
colRemoveidx = c();

for (i in 1:colCnt)
  if(is.factor(training[[i]]))
    colRemoveidx <- c(colRemoveidx, i)

training <- training[,-colRemoveidx]

training <- training[,-c(9:24, 38:47, 57:65, 69:74,76:85,98:114)]
training <- training[,-c(1:4)]
```



###Create data partitions on training set
Keeping the original test-data aside, we partition the training data into training and test. 
```{r, echo=TRUE}

inTrain <- createDataPartition(y=training$classe, p = 0.7, list = F)
subTrain <- training[inTrain,]
subTest <- training[-inTrain,]

```

###Apply Random Forest algo to generate model
```{r, echo=TRUE}

totalPredictorCount <- dim(subTest)[2]
fit <- randomForest(classe ~. ,data=subTrain,mtry=sqrt(totalPredictorCount),importance=TRUE)


```

###Out of sample error/Confusion matrix
```{r, echo=TRUE}

print(fit)

```


###Cross-validation on training's test set.
```{r, echo=TRUE}
pred <- predict(fit, newdata=subTest)
confusionMatrix(subTest$classe, pred)
```


